Floquet-Drude Conductivity in Dressed Quantum Hall Systems

### 2021-07-23 paper/v1.0/sections/1_introduction.tex
Formatting and corrections

Theoretical analyses of the transport properties of dressed fermions systems were recently reported in Ref. \cite{kibis14,morina15,wackerl20} where a general expression for Floquet-Drude conductivity has been derived in a fully closed analytical form in Ref. \cite{wackerl20}, introducing a novel type of Green’s functions namely four-times Green’s functions. As a consequence, the established formalism introduces a novel technique to exploit the transport attributes of nanostructures by an intense dressing field.

@@

In contrast, more fascinating transport phenomena can be achieved by creating a quantum Hall system using a nonoscillating magnetic field and simultaneously applying a dressing field.
Whilst there exist a number of leading theories devoted to the calculation of conductivity tensor elements in quantum Hall systems \cite{ando74_1,ando82,endo09}, they have not been utilized to describe the optical manipulation of charge transport in the quantum Hall systems.
Furthermore, K. Dini \textit{et al.} \cite{dini16} have recently investigated the one-directional conductivity behavior of dressed quantum Hall systems, and they have not adopted the state-of-the-art model to describe the conductivity in a quantum Hall system. In their study, they used the conductivity models from Ref. \cite{ando74_1,ando82} and as mentioned in A. Endo \textit{et al.} work \cite{endo09} those models yields a semi-elliptical broadening for Landau levels and gives barely fitting representations for the experimental findings.

>>>>

Theoretical analyses of the transport properties of dressed fermions systems were recently reported in Refs. \cite{kibis14,morina15,wackerl20} where a general expression for Floquet-Drude conductivity has been derived in a fully closed analytical form in Ref. \cite{wackerl20}, introducing a novel type of Green’s functions namely four-times Green’s functions. As a consequence, the established formalism introduces a novel technique to exploit the transport attributes of nanostructures by an intense dressing field.

@@

In contrast, more fascinating transport phenomena can be achieved by creating a quantum Hall system using a nonoscillating magnetic field and simultaneously applying a dressing field.
Whilst there exist a number of leading theories devoted to the calculation of conductivity tensor elements in quantum Hall systems \cite{ando74_1,ando82,endo09}, they have not been utilized to describe the optical manipulation of charge transport in the quantum Hall systems.
Furthermore, K. Dini \textit{et al.} \cite{dini16} have recently investigated the one-directional conductivity behavior of dressed quantum Hall systems, and they have not adopted the state-of-the-art model to describe the conductivity in a quantum Hall system. In their study, they used the conductivity models from Refs. \cite{ando74_1,ando82} and as mentioned in A. Endo \textit{et al.} work \cite{endo09} those models yields a semi-elliptical broadening for Landau levels and gives barely fitting representations for the experimental findings.

### 2021-07-23 paper/v1.0/sections/2_schrodinger_problem.tex
Formatting and corrections

Using Landau gauge for the stationary magnetic field, we can represent it using a vector potential as $\vb{A}_{s} = (-By,0,0)^{\text{T}}$ and choosing Coulomb gauge, the vector potential of the dynamic dressing radiation can be presented as $\vb{A}_{d}(t) = (0,[E/\omega ]\cos(\omega t),0)^{\text{T}}$. These vector potentials are coupled to the momentum of 2DEG as kinetic momentum \cite{mahan00,bruus04} and this leads to the time-dependent Hamiltonian
\begin{equation} \label{eq_1}
  \hat{H}_e(t) = \frac{1}{2m_e}\Big[\hat{\vb{p}} - e\big[\vb{A}_{s}+\vb{A}_{d}(t)\big]\Big]^2,
\end{equation}
where $m_e$ is the effective electron mass and $e$ is the magnitude of the electron charge. $\hat{\vb{p}} = (\hat{p}_x,\hat{p}_y,0)^{\text{T}}$ represents the canonical momentum operator for 2DEG with electron momentum $p_{x,y}$.
The exact solutions for the time-dependent Schrödinger equation $i\hbar \dv{t}\psi = \hat{H}_e(t) \psi$ was already given by Refs. \cite{husimi53,ditt98,dini16} and we can present them as a set of wave functions defined by two quantum numbers $(n,m)$
\begin{equation} \label{eq_2}
  \begin{aligned}
    \psi_{n,m}&(x,y,t)  \\
    & = \frac{1}{\sqrt{L_x}}
    \chi_n\big(y - y_0 - \zeta(t)\big)\\
    & \times
    \text{exp}\bigg(
    \frac{i}{\hbar}\bigg[- \varepsilon_nt
    + p_x x + \frac{eE[y - y_0]}{\omega}\cos(\omega t)\\
    &+
    m_e\dot{\zeta}(t)\big[y - y_0 -\zeta(t)\big] +
    \int_0^{t}dt'L(\zeta,\dot{\zeta},t')\bigg]\bigg),
  \end{aligned}
\end{equation}
where $n \in \mathbb{Z}^{+}_0$ and $m \in \mathbb{Z}$ ; see Appendix A. Here $L_{x,y}$ are dimension of the 2DEG surface, $\hbar$ is the reduced Planck constant, and the center of the cyclotron orbit on the $y$-axis is given by $y_0 \equiv -p_x/eB$ with $p_x = 2\pi \hbar m/L_x$. $\chi_n$ are well known solutions for Schrödinger equation of a stationary quantum harmonic oscillator
\begin{equation} \label{eq_3}
  \chi_n(y) =
   \frac{\sqrt{\kappa}}{\sqrt{2^{n}n!}}
  e^{-\kappa^2 y^2/2}
  \mathcal{H}_n \qty(\kappa y) \quad \text{with}
  \quad
  \kappa = \sqrt{\frac{m_e \omega_0}{\hbar}},
\end{equation}
where eigenvalues are given by $\varepsilon_n = \hbar \omega_0 (n + 1/2)$ and $\omega_0 \equiv eB/m_e$ is the cyclotron frequency. Each $n$ value defines the occupied energy($\varepsilon_n$) of the respective Landau level. The path shift of the driven classical oscillator $\zeta(t)$ is given by
\begin{equation} \label{eq_4}
  \zeta(t) = \frac{eE}{m_e(\omega_0^2 - \omega^2)}\sin(\omega t),
\end{equation}
while the Lagrangian of the classical oscillator $L(\zeta,\dot{\zeta},t)$ can be identified as
\begin{equation} \label{eq_5}
  L(\zeta,\dot{\zeta},t) = \frac{1}{2} m_e\dot{\zeta}^2(t) - \frac{1}{2}m_e\omega_0^2 \zeta^2(t) + eE\zeta(t) \sin(\omega t).
\end{equation}
The exponential phase shifts in Eq.~(\ref{eq_2}) represent the influence done by the stationary magnetic field and dressing field. Therefore, we can accept that magneto-transport characteristics of 2DEG can be renormalized by the nonoscillating magnetic field along with the dressing field.

>>>>

Using Landau gauge for the stationary magnetic field, we can represent it using a vector potential as $\vb{A}_{s} = (-By,0,0)^{\text{T}}$ and choosing Coulomb gauge, the vector potential of the dynamic dressing radiation can be presented as $\vb{A}_{d}(t) = (0,[E/\omega ]\cos(\omega t),0)^{\text{T}}$. These vector potentials are coupled to the momentum of 2DEG as kinetic momentum \cite{mahan00,bruus04} and this leads to the time-dependent Hamiltonian
\begin{equation} \label{eq_1}
  \hat{H}_e(t) = \frac{1}{2m_e}\Big[\hat{\vb{p}} - e\big[\vb{A}_{s}+\vb{A}_{d}(t)\big]\Big]^2,
\end{equation}
where $m_e$ is the effective electron mass and $e$ is the magnitude of the electron charge. $\hat{\vb{p}} = (\hat{p}_x,\hat{p}_y,0)^{\text{T}}$ represents the canonical momentum operator for 2DEG with electron momentum $p_{x,y}$.
The exact solutions for the time-dependent Schrödinger equation $i\hbar \dv{t}\psi = \hat{H}_e(t) \psi$ was already given by Refs. \cite{husimi53,ditt98,dini16} and we can present them as a set of wave functions defined by two quantum numbers $(n,m)$
\begin{equation} \label{eq_2}
  \begin{aligned}
    \psi_{n,m}&(x,y,t)  \\
    & = \frac{1}{\sqrt{L_x}}
    \chi_n\big(y - y_0 - \zeta(t)\big)\\
    & \times
    \text{exp}\bigg(
    \frac{i}{\hbar}\bigg[- \epsilon_n t
    + p_x x + \frac{eE[y - y_0]}{\omega}\cos(\omega t)\\
    &+
    m_e\dot{\zeta}(t)\big[y - y_0 -\zeta(t)\big] +
    \int_0^{t}dt'L(\zeta,\dot{\zeta},t')\bigg]\bigg),
  \end{aligned}
\end{equation}
where $n \in \mathbb{Z}^{+}_0$ and $m \in \mathbb{Z}$ ; see Appendix \ref{appendix_a}. Here $L_{x,y}$ are dimension of the 2DEG surface, $\hbar$ is the reduced Planck constant, and the center of the cyclotron orbit on the $y$-axis is given by $y_0 \equiv -p_x/eB$ with $p_x = 2\pi \hbar m/L_x$. $\chi_n$ are well known solutions for Schrödinger equation of a stationary quantum harmonic oscillator
\begin{equation} \label{eq_3}
  \chi_n(y) =
   \frac{\sqrt{\kappa}}{\sqrt{2^{n}n!}}
  e^{-\kappa^2 y^2/2}
  \mathcal{H}_n \qty(\kappa y) \quad \text{with}
  \quad
  \kappa = \sqrt{\frac{m_e \omega_0}{\hbar}},
\end{equation}
where eigenvalues are given by $\epsilon_n = \hbar \omega_0 (n + 1/2)$ and $\omega_0 \equiv eB/m_e$ is the cyclotron frequency. Each $n$ value defines the occupied energy($\epsilon_n$) of the respective Landau level. The path shift of the driven classical oscillator $\zeta(t)$ is given by
\begin{equation} \label{eq_4}
  \zeta(t) = \frac{eE}{m_e(\omega_0^2 - \omega^2)}\sin(\omega t),
\end{equation}
while the Lagrangian of the classical oscillator $L(\zeta,\dot{\zeta},t)$ can be identified as
\begin{equation} \label{eq_5}
  L(\zeta,\dot{\zeta},t) = \frac{1}{2} m_e\dot{\zeta}^2(t) - \frac{1}{2}m_e\omega_0^2 \zeta^2(t) + eE\zeta(t) \sin(\omega t).
\end{equation}
The exponential phase shifts in Eq.~(\ref{eq_2}) represent the influence done by the stationary magnetic field and dressing field. Therefore, we can accept that magneto-transport characteristics of 2DEG can be renormalized by the nonoscillating magnetic field along with the dressing field.

### 2021-07-23 paper/v1.0/sections/4_inverse_scattering_time.tex
Formatting and corrections

In Ref. \cite{wackerl20} Floquet Fermi golden rule has been introduced as a method to analyze transport properties in dressed disordered quantum systems. However, this theory has not been applied for a dressed quantum Hall system and to identify magneto-transport properties in our system we use Floquet Fermi golden rule.
With the help of $t-t'$ formalism \cite{wackerl20,grifoni98,sambe75,peskin93,althorpe97} and using Floquet states derived in Eq.~(\ref{eq_12}) we can derive an  expression for the inverse scattering time matrix ($(l,l')$th element) for the $N$-th Landau level, per a given energy $\varepsilon$ and momentum $k_x$ value for our considered quantum Hall system as
\begin{equation} \label{eq_13}
  \begin{aligned}
    \bigg(&\frac{1}{\tau(\varepsilon,k_x)}\bigg)^{ll'}_{N} \\
    & =
    \frac { \varrho^2}{eB}
    \delta(\varepsilon - \varepsilon_{N}) \\
    & \times
    \int_{-\infty}^{\infty} d k_1 \Bigg[
    J_l\qty(\frac{b\hbar}{eB}[{k}_x - k_1])
    J_{l'}\qty(\frac{b\hbar}{eB}[{k}_x - k_1]) \\
    & \times
    \qty|
    \int_{-\infty}^{\infty} dk_2 \;
    {\chi}_{N}\qty(\frac{\hbar}{eB}k_2)
    {\chi}_{N}\qty(\frac{\hbar}{eB} \qty[k_1 - {k}_x - k_2])|^2\Bigg],
  \end{aligned}
\end{equation}
where $\varrho \equiv \eta_{imp} L_x [ { V_{imp}}/{eB}]^{1/2}$,
$J_l(\cdot)$ are Bessel functions of the first kind with $l$-th integer order and $\varepsilon_N$ is the energy of $N$-th Landau level; see Appendix C. In this derivation, we assumed that the perturbation potential is generated by a group of randomly distributed impurities, since random impurities in a disordered metal is a better approximation for experimental results. Thus, the total scattering potential in 2DEG has been presented as the sum of uncorrelated single impurity potentials $\upsilon(\vb{r})$. Here $\eta_{imp}$ is the number of impurities in a unit area and $V_{imp} \equiv \expval{|V_{{k'}_x,k_x}|^2}$ with $V_{{k'}_x,k_x} = \mel**{k'_x}{\upsilon(x) }{k_x}$ where $\braket{x}{k_x} = e^{-ik_x x}/\sqrt{L_x}$.

>>>>

In Ref. \cite{wackerl20} Floquet Fermi golden rule has been introduced as a method to analyze transport properties in dressed disordered quantum systems. However, this theory has not been applied for a dressed quantum Hall system and to identify magneto-transport properties in our system we use Floquet Fermi golden rule.
With the help of $t-t'$ formalism \cite{wackerl20,grifoni98,sambe75,peskin93,althorpe97} and using Floquet states derived in Eq.~(\ref{eq_12}) we can derive an  expression for the inverse scattering time matrix ($(l,l')$th element) for the $N$-th Landau level, per a given energy $\varepsilon$ and momentum $k_x$ value for our considered quantum Hall system as
\begin{equation} \label{eq_13}
  \begin{aligned}
    \bigg(&\frac{1}{\tau(\varepsilon,k_x)}\bigg)^{ll'}_{N} \\
    & =
    \frac { \varrho^2}{eB}
    \delta(\varepsilon - \varepsilon_{N}) \\
    & \times
    \int_{-\infty}^{\infty} d k_1 \Bigg[
    J_l\qty(\frac{b\hbar}{eB}[{k}_x - k_1])
    J_{l'}\qty(\frac{b\hbar}{eB}[{k}_x - k_1]) \\
    & \times
    \qty|
    \int_{-\infty}^{\infty} dk_2 \;
    {\chi}_{N}\qty(\frac{\hbar}{eB}k_2)
    {\chi}_{N}\qty(\frac{\hbar}{eB} \qty[k_1 - {k}_x - k_2])|^2\Bigg],
  \end{aligned}
\end{equation}
where $\varrho \equiv \eta_{imp} L_x [ { V_{imp}}/{eB}]^{1/2}$,
$J_l(\cdot)$ are Bessel functions of the first kind with $l$-th integer order and $\varepsilon_N$ is the energy of $N$-th Landau level; see Appendix \ref{appendix_c}. In this derivation, we assumed that the perturbation potential is generated by a group of randomly distributed impurities, since random impurities in a disordered metal is a better approximation for experimental results.
Thus, the total scattering potential in 2DEG has been presented as the sum of uncorrelated single impurity potentials $\upsilon(\vb{r})$. Here $\eta_{imp}$ is the number of impurities in a unit area and $V_{imp} \equiv \expval{|V_{{k'}_x,k_x}|^2}_{imp}$ with $V_{{k'}_x,k_x} = \mel**{k'_x}{\upsilon(x) }{k_x}$ where $\braket{x}{k_x} = e^{-ik_x x}$.
Here $\expval{\cdot}_{imp}$ represents the average over the impurity disorder.

### 2021-07-23 paper/v1.0/sections/5_floquet_drude_conductivity.tex
Formatting and corrections

A general theory for the conductivity in dressed systems with the disorder averaging was introduced by M. Wackerl in Ref.~\cite{wackerl20,wackerlthesis20}. Within this theory, the general $x$-directional longitudinal DC limit conductivity can be express with
\begin{equation} \label{eq_20}
  \begin{aligned}
    {\sigma}^{xx} &=
    \frac{-1}{4\pi\hbar A}
    \int_{\Pi-\hbar\omega/2}^{\Pi+ \hbar\omega/2} d\varepsilon \bigg[
    \qty(
    -\frac{\partial f}{\partial \varepsilon})
    \\
    & \times
    \tr
    \qty[
    {j}^x_0
    \qty(
    \vb{G}^{r} (\varepsilon) - \vb{G}^{a} (\varepsilon)
    )
    {j}^x_0
    \qty(
    \vb{G}^{r} (\varepsilon) - \vb{G}^{a} (\varepsilon)
    )
    ]\bigg],
  \end{aligned}
\end{equation}
where ${j}^x_0$ and $\vb{G}^{r,a} (\varepsilon)$ are $x$-directional electric current operator matrix elements' $s=0$ the Fourier component (see Appendix D) and white noise disorder averaged Floquet Green function matrix \cite{wackerl20,wackerlthesis20} respectively defined against the Floquet modes of the considering system. Here we have assumed that only $s=0$ Fourier component of the current operator is contributing to the conductivity. In addition, $A$ is the area of the considering two-dimentional system, partial distribution function is given by $f$ and $\Pi$ is a function that can be chosen such that
\begin{equation} \label{eq_21}
    \Pi- \frac{\hbar \omega}{2}
    \leq \varepsilon_N
    <
    \Pi + \frac{\hbar \omega}{2},
\end{equation}
where $ \varepsilon_N$ are quasienergies of all considering Floquet states.

@@

Since we can diagonalize the impurity averaged Green's function using unitary transformation ($\vb{T}  \equiv \ket{n,k_x}$) mentioned in Ref.~\cite{wackerl20,wackerlthesis20,tsuji08}, and we evaluate the matrix element of the difference between retarded and advanced Green's functions

@@

using the retarded self-energy matrix $\vb{\Sigma}^r$ which is the sum of all irreducible diagrams \cite{wackerl20,wackerlthesis20}. Using the matrix elements of electric current operator in Landau levels (see Appendix D) and

@@

Same as above derivation we can derive the longitudinal conductivity in $y$-direction by using the current operator derived in Appendix D

>>>>

A general theory for the conductivity in dressed systems with the disorder averaging was introduced by M. Wackerl in Refs.~\cite{wackerl20,wackerlthesis20}. Within this theory, the general $x$-directional longitudinal DC limit conductivity can be express with
\begin{equation} \label{eq_20}
  \begin{aligned}
    {\sigma}^{xx} &=
    \frac{-1}{4\pi\hbar A}
    \int_{\Pi-\hbar\omega/2}^{\Pi+ \hbar\omega/2} d\varepsilon \bigg[
    \qty(
    -\frac{\partial f}{\partial \varepsilon})
    \\
    & \times
    \tr
    \qty[
    {j}^x_0
    \qty(
    \vb{G}^{r} (\varepsilon) - \vb{G}^{a} (\varepsilon)
    )
    {j}^x_0
    \qty(
    \vb{G}^{r} (\varepsilon) - \vb{G}^{a} (\varepsilon)
    )
    ]\bigg],
  \end{aligned}
\end{equation}
where ${j}^x_0$ and $\vb{G}^{r,a} (\varepsilon)$ are $x$-directional electric current operator matrix elements' $s=0$ the Fourier component (see Appendix \ref{appendix_d}) and white noise disorder averaged Floquet Green function matrix \cite{wackerl20,wackerlthesis20} respectively defined against the Floquet modes of the considering system. Here we have assumed that only $s=0$ Fourier component of the current operator is contributing to the conductivity. In addition, $A$ is the area of the considering two-dimentional system, partial distribution function is given by $f$ and $\Pi$ is a function that can be chosen such that
\begin{equation} \label{eq_21}
    \Pi- \frac{\hbar \omega}{2}
    \leq \varepsilon_N
    <
    \Pi + \frac{\hbar \omega}{2},
\end{equation}
where $ \varepsilon_N$ are quasienergies of all considering Floquet states.

@@

Since we can diagonalize the impurity averaged Green's function using unitary transformation ($\vb{T}  \equiv \ket{n,k_x}$) mentioned in Refs.~\cite{wackerl20,wackerlthesis20,tsuji08}, and we evaluate the matrix element of the difference between retarded and advanced Green's functions

@@

using the retarded self-energy matrix $\vb{\Sigma}^r$ which is the sum of all irreducible diagrams \cite{wackerl20,wackerlthesis20}. Using the matrix elements of electric current operator in Landau levels (see Appendix \ref{appendix_d}) and

@@

Same as above derivation we can derive the longitudinal conductivity in $y$-direction by using the current operator derived in Appendix \ref{appendix_d}

### 2021-07-23 paper/v1.0/sections/6_manipulate_conductivity.tex
Formatting and corrections

Considering the effect of the external dressing field on longitudinal conductivity of 2DEG, we can identify that high intensities shrink the conductivity regions near Landau levels. However, the peak value of the conductivity at each Landau level has the same value as the undressed system. This demonstrate that we are able to tune the width of the regions of conductivity in these quantum Hall systems with the help of a dressing field.
These characteristics are aligned with results demonstrated by K. Dini \textit{et al.} \cite{dini16} and as they remarked since the Fermi level of the system can be change with the applied gate voltage of the material this can be used as a 2D switch for nanoscale optoelectronics applications. Controlling  the external dressing field we are able to fine-tune the switching mechanism for optimized performance.
Furthermore, we can distinguish that the shapes and behavior of the conductivity regions illustrated in Fig.~\ref{fig_5} and \ref{fig_6} are generally incompatible with the results reported in Ref.~\cite{dini16}. This is due to the selection of the conventional longitudinal conductivity theory of 2DEG from Ref.~\cite{ando74_1,ando82}. The semi-elliptical conductivity regions illustrated in Ref.~\cite{dini16,ando74_1,ando82}, have less consistence with the experimentally observed Landau levels representation \cite{endo09}.
In our study on the transport properties of quantum Hall systems, we developed the conductivity expression starting from Floquet-Drude conductivity \cite{wackerl20} and our results are much more align with the result mentioned in Ref. \cite{endo09}.
The description of conductivity of quantum Hall systems demonstrated in Ref.~\cite{endo09} has excellent agreement between the theory and experiment obtained in a GaAs/AlGaAs 2DES for the low magnetic field range. However, they have not considered the tunability that can be achieved with the external strong dressing field. In this analysis we account both magnetic and dressing field effects that can be applied into the transport properties of 2DEG, and we have presented a more generalized theory. As a concluding remark, in this study we were able to demonstrate that using Floquet-Drude conductivity method one can derive a more experimental fitting and generalized mathematical model that describes the transport properties of quantum Hall systems.

>>>>

Considering the effect of the external dressing field on longitudinal conductivity of 2DEG, we can identify that high intensities shrink the conductivity regions near Landau levels. However, the peak value of the conductivity at each Landau level has the same value as the undressed system. This demonstrate that we are able to tune the width of the regions of conductivity in these quantum Hall systems with the help of a dressing field.
These characteristics are aligned with results demonstrated by K. Dini \textit{et al.} \cite{dini16} and as they remarked since the Fermi level of the system can be change with the applied gate voltage of the material this can be used as a 2D switch for nanoscale optoelectronics applications. Controlling  the external dressing field we are able to fine-tune the switching mechanism for optimized performance.
Furthermore, we can distinguish that the shapes and behavior of the conductivity regions illustrated in Fig.~\ref{fig_5} and \ref{fig_6} are generally incompatible with the results reported in Ref.~\cite{dini16}. This is due to the selection of the conventional longitudinal conductivity theory of 2DEG from Refs.~\cite{ando74_1,ando82}. The semi-elliptical conductivity regions illustrated in Refs.~\cite{dini16,ando74_1,ando82}, have less consistence with the experimentally observed Landau levels representation \cite{endo09}.
In our study on the transport properties of quantum Hall systems, we developed the conductivity expression starting from Floquet-Drude conductivity \cite{wackerl20} and our results are much more align with the result mentioned in Ref. \cite{endo09}.
The description of conductivity of quantum Hall systems demonstrated in Ref.~\cite{endo09} has excellent agreement between the theory and experiment obtained in a GaAs/AlGaAs 2DES for the low magnetic field range. However, they have not considered the tunability that can be achieved with the external strong dressing field. In this analysis we account both magnetic and dressing field effects that can be applied into the transport properties of 2DEG, and we have presented a more generalized theory. As a concluding remark, in this study we were able to demonstrate that using Floquet-Drude conductivity method one can derive a more experimental fitting and generalized mathematical model that describes the transport properties of quantum Hall systems.
